The Hyperedge Event Model

### 2018-09-09 paper/revision1/ba-sample.tex
revised notations

	The hyperedge event model (HEM) specifies a generative process for $E$ unique hyperedge events that occur between $A$ nodes. A single hyperedge event is indexed by $e \in [E]$---where $[E]$ denotes a categorical set with $E$ levels $[E] = \{1,\ldots,E\}$---and consists of three components: the sender $s_e \in [A]$, an indicator vector of receivers $\boldsymbol{r}_e$---where $r_{ej}=1$ if $j \in [A]$ is a receiver of hyperedge event $e$ and 0 otherwise---and the timestamp $t_e \in (0, \infty)$. For simplicity, we assume that events are ordered by time such that $t_e \leq t_{e+1}$. While the model can be applied to two types of hyperedge events---events involving (1) one sender and one or more receivers, and (2) one or more senders and one receiver---here we only present the generative process for those involving one sender and one or more receivers (i.e., multicast). One notable feature of our generative process is that we draw auxiliary variables that serve as candidate data. Data is generated from the HEM through a sampling process applied to the auxiliary variables. The auxiliary variables drawn for event $e$ include, for each possible sender $i \in [A]$, a time increment from event $e-1$ at which sender $i$ would create event $e$, and an $A-1$ length vector indicating which nodes would be the receivers of event $e$ if it were directed by sender $i$.  The data generated for event $e$ under the HEM corresponds to the sender that would create event $e$ the soonest---at the smallest time increment from event $e-1$. The receivers of event $e$ generated under the HEM correspond to those receivers toward which the sender with the minimum time increment would have directed event $e$. We explain these steps in more detail below. Figure \ref{figure:diagram} presents an illustrative example on how the $e^{\textrm{th}}$ event is generated, assuming $t_{e-1} = 0$ and $A=5$. For hyperedge events that involve one receiver and one or more senders, we treat $s_e$ to be an indicator vector of senders $\boldsymbol{s}_e$ and $r_e$ to be the single receiver, and then follow the alternative generative process provided in Appendix A, which we derive as a simple reversal of the process used for multicasts (i.e., one sender and multiple receivers).~\looseness=-1 

@@

	The HEM assumes that the sender of each hyperedge event is the sender that would initiate their respective event with the greatest urgency (i.e., the earliest timestamp). Our model thus assumes that for every event $e$, every possible sender $i$ generates a candidate receiver set that would be the receiver set of event $e$ if sender $i$ were the sender. For an event $e$, we first define an $A\times A$ matrix $\boldsymbol{u}_e$ where the $i^{\textrm{th}}$ row denotes sender $i$'s receiver vector of zeros and 1's---i.e., 1's indicate the nodes to which $i$ intends to direct event $e$. We then assume that each receiver vector $\boldsymbol{u}_{ie}$ comes from a modification of the multivariate Bernoulli (MB) distribution \citep{dai2013multivariate}---a model that has been used to model graphs in which the state of each edge indicator is drawn independently from an edge-specific Bernoulli distribution. Motivated by the Gibbs measure \citep{fellows2017removing}, we define a probability measure ``MB$_{G}$" in order to avoid drawing hyperedge events with no receiver. This probability measure amounts to a non-empty Gibbs measure, in which the all-zero vector is excluded from the support of the multivariate Bernoulli distribution. As a result, this measure helps us to (1) allow a sender to select multiple receivers for a single event, (2) force the sender to select at least one receiver, and (3) ensure a tractable normalizing constant for the receiver selection distribution. To be specific, we draw a binary vector $\boldsymbol{u}_{ie}= (u_{ie1},
	\ldots, u_{ieA})$ for $i$th row of matrix $\boldsymbol{u}$ using

@@

	\ldots, \lambda_{ieA})$ and $\lambda_{iej}$ is the linear combination of statistics relevant to the receiver selection process for every possible sender--receiver pair $(i,j)$ where $i \!\neq\! j$. In other words, we define

@@

	In particular, we define $\mbox{MB}_{G}(\boldsymbol{\lambda}_{ie})$ as
	\begin{equation}
		\begin{aligned}
			&\Pr(\boldsymbol{u}_{ie}|\,\boldsymbol{b}, \boldsymbol{x}_{ie}) = \frac{1}{Z(\boldsymbol{\lambda}_{ie})}\exp\Big(\mbox{log}\big(\text{I}( \lVert \boldsymbol{u}_{ie}\rVert_1 > 0 )\big) + \sum_{j\neq i} \lambda_{iej}u_{iej}\Big) ,
		\end{aligned}
		\label{eqn:Gibbs}
	\end{equation}
	where $Z(\boldsymbol{\lambda}_{ie})$ is the normalizing constant, $\lVert \cdot \rVert_1$ is the $\ell_1$-norm, and the log-indicator term $\mbox{log}\big(\text{I}( \lVert \boldsymbol{u}_{ie}\rVert_1 > 0 )\big)$ ensures that empty receiver sets are excluded from the distribution's support. These modeling assumptions facilitate efficient posterior inference since we can derive a closed form expression for the normalizing constant---i.e., $Z(\boldsymbol{\lambda}_{ie})= \prod_{j\neq i} \big(\mbox{exp}(\lambda_{iej}) + 1\big)-1$---and thus do not need to perform brute-force summation over the support of $\boldsymbol{u}_{ie} \in \{0,1\}^A$. We provide detailed derivation steps for the normalizing constant $Z(\boldsymbol{\lambda}_{ie})$ in Appendix B.

	% $Z(\boldsymbol{\lambda}_{ie})= \prod_{j\neq i} \big(\mbox{exp}(\lambda_{iej}) + 1\big)-1$

	\subsection{Candidate timestamps}\label{subsec:Time}	
	 In modeling ``when," we do not directly model the timestamp $t_e$. Instead, we assume that each sender's ``time increment"---i.e., waiting time to next event since $t_{e-1}$---is drawn from a specific exponential family distribution. We assume that every sender $i \in [A]$ generates his/her own time increment from event $e-1$ to event $e$, which we define as $\tau_{ie}$. We further specify that the distribution of $\tau_{ie}$ has sender-specific mean $\mu_{ie}$ and common variance $V(\mu)$. Following the generalized linear model (GLM) framework \citep{nelder1972generalized}, we assume that sender $i$'s time increment $\tau_{ie}$ satisfy~\looseness=-1
	\begin{equation}
		\begin{aligned}
			E(\tau_{ie}) &= \mu_{ie},\\
			V(\tau_{ie}) &= V(\mu),
		\end{aligned}
	\end{equation}
		% \footnote{The log-normal distribution is not exponential family but can be used via modeling of $\log(\tau_e)$.} 
	where $\tau_{ie}$ here is a positive real number. Possible choices of distribution include exponential, Weibull, gamma, and log-normal distributions, which are commonly used in time-to-event modeling \citep{rao2000applied,rizopoulos2012joint}. Based on the specific distribution, we may need other latent variables to draw the time increment, to account for the variance of time increments $V(\mu)$---e.g., the shape parameter $k$ for the Weibull, the shape parameter $\theta$ for the gamma, and the variance parameter $\sigma_\tau^2$ for the log-normal---beyond the coefficients for the features used to model the rate. We use $f_\tau(\cdot; \mu, V(\mu))$ and $F_\tau(\cdot; \mu, V(\mu))$ to denote the probability density function (p.d.f) and cumulative density function (c.d.f), respectively, with mean $\mu$ and variance $V(\mu)$. Similar to Section \ref{subsec: Tie}, the timing rate $\mu_{ie}$ for sender $i$ and event $e$ follows
	\begin{equation}
	\mu_{ie} = g^{-1}(\boldsymbol{\eta}^\top \boldsymbol{y}_{ie}),
	\end{equation}
	where $\boldsymbol{\eta}$ is a $Q$-dimensional vector of coefficients with a Normal prior $\boldsymbol{\eta} \sim N(\boldsymbol{\mu}_\eta,\Sigma_\eta)$, $\boldsymbol{y}_{ad}$ is a set of event timing features---covariates that could affect timestamps of events, and $g(\cdot)$ is the appropriate link function such as identity, log, or inverse.

	\subsection{Senders, receivers, and timestamps}\label{subsec:Observed}
		After the HEM draws the candidate timestamp vector $\boldsymbol{\tau}_e=\{\tau_{1e},\ldots,\tau_{Ae}\}$ at which the event would be created given the candidate sender and receiver combinations, finally our model assumes that the observed sender, receivers, and timestamp of hyperedge event $e$ are generated by selecting the sender--receiver-set pair with the smallest time increment:
	\begin{equation}
		\begin{aligned}
			s_e &= \mbox{argmin}_{i}(\tau_{ie}),\\
			\boldsymbol{r}_e &= \boldsymbol{u}_{s_e e},\\
			t_e &=t_{e-1} + \tau_{s_e e}.
		\end{aligned}
	\end{equation}
	\begin{algorithm}[!t]
		\spacingset{1}
		\SetAlgoLined
		\caption{Generative process: one sender and one or more receivers}
		\begin{algorithmic}
			\STATE \textbf{Input}: number of events and nodes $(E, A)$, covariates $(\boldsymbol{x}, \boldsymbol{y})$, and coefficients $(\boldsymbol{b}, \boldsymbol{\eta})$
			\vskip 0.1in
			\FOR{$e=1$ to $E$}
			\FOR{$i=1$ to $A$}
			\FOR{$j=1$ to $A$ ($j \neq i$)}
			\STATE	set $\lambda_{iej} = {\boldsymbol{b}}^{\top}\boldsymbol{x}_{iej}$
			\ENDFOR
			\STATE	draw $\boldsymbol{u}_{ie}  \sim
			\mbox{MB}_G(\boldsymbol{\lambda}_{ie})$
			\STATE		set $\mu_{ie} = g^{-1}(\boldsymbol{\eta}^\top \boldsymbol{y}_{ie})$
			\STATE		draw $\tau_{ie} \sim f_\tau(\mu_{ie}, V(\mu))$
			\ENDFOR

			\IF{$n \geq 2$ tied events} 
			\STATE	set $s_e,\ldots, s_{e+n-1}=\mbox{argmin}_{i}(\tau_{ie}),$
			\STATE	set $\boldsymbol{r}_e=\boldsymbol{u}_{s_e e},\ldots,\boldsymbol{r}_{e+n-1}=\boldsymbol{u}_{s_{e+n-1} e}$
			\STATE	set $t_e, \ldots, t_{e+n-1}=t_{e-1} + \min_i\tau_{ie}$
			\STATE		jump to $e = e+n$
			\ELSE
			\STATE	set $s_e= \mbox{argmin}_{i}(\tau_{ie})$
			\STATE		set $\boldsymbol{r}_e = \boldsymbol{u}_{s_e e}$
			\STATE	set $t_e =t_{e-1} + \min_i\tau_{ie}$
			\ENDIF
			\ENDFOR
		\end{algorithmic}
		\label{alg:generative}
	\end{algorithm}
		Therefore, the HEM assumes a sender-driven process---i.e., the receivers and timestamp of an event are jointly determined by the sender's urgency to direct the event to those selected receivers---which is similar to \cite{snijders1996stochastic} in that each actor gets his/her own rate function to determine ``who will make the next change in network." Note that our generative process allows for tied events. In the case of tied events---i.e., multiple senders draw exactly the same candidate timestamps---we assume that all events are generated and occur simultaneously. Algorithm \ref{alg:generative} summarizes the entire generative process for hyperedge events with one sender and one or more receivers.~\looseness=-1

	\section{Posterior inference}\label{sec:inference}
	In this section we describe how we invert the generative process to obtain the posterior distribution over the latent variables---candidate receivers $\{\boldsymbol{u}_e\}_{e=1}^E$, coefficients for receiver selection features $\boldsymbol{b}$, and coefficients for event timing features $\boldsymbol{\eta}$---conditioned on the observed data $\{(s_e, \boldsymbol{r}_e, t_e)\}_{e=1}^E$, covariates $\{(\boldsymbol{x}_e, \boldsymbol{y}_e)\}_{e=1}^E$, and hyperparameters $(\boldsymbol{\mu}_b, \Sigma_b, \boldsymbol{\mu}_\eta, \Sigma_\eta)$. We draw the samples using Markov chain Monte Carlo (MCMC) methods, repeatedly resampling the value of each latent variable from its conditional posterior via a Metropolis-within-Gibbs sampling algorithm. In the next subsection, we provide each latent variable's conditional posterior along with pseudocode of MCMC in Algorithm \ref{alg:MCMC}. We also evaluate the correctness of both our mathematical derivations and software implemenation using the prior--posterior simulator test of \cite{geweke2004getting} in Appendix C.

@@

	In our model, direct computation of the posterior densities for the latent variables $\boldsymbol{b}$ and $\boldsymbol{\eta}$---i.e., $\Pr(\boldsymbol{b}|\,\boldsymbol{x},\boldsymbol{s}, \boldsymbol{r},\boldsymbol{t})$ and $\Pr(\boldsymbol{\eta}|\,\boldsymbol{y},\boldsymbol{s}, \boldsymbol{r},\boldsymbol{t})$---is not possible. However, it is possible to augment the data by candidate receivers $\boldsymbol{u}$ such that we can obtain their conditional posterior by conditioning on samples of $\boldsymbol{u}$. %from $P(\boldsymbol{b}, \boldsymbol{u}|\,\boldsymbol{x},\boldsymbol{a}, \boldsymbol{r},\boldsymbol{t})$ and $P(\boldsymbol{\eta}, \boldsymbol{u}|\, \boldsymbol{y},\boldsymbol{a}, \boldsymbol{r},\boldsymbol{t})$.

@@

		\prod_{i=1}^A \frac{1}{Z(\boldsymbol{\lambda}_{ie})}\exp\Big(\mbox{log}\big(\text{I}( \lVert \boldsymbol{u}_{ie}\rVert_1 > 0)\big) + \sum\limits_{j \neq i} \lambda_{iej}u_{iej}\Big).

@@

		\Pr(\boldsymbol{\eta}|\, \boldsymbol{u}, \boldsymbol{y},\boldsymbol{s}, \boldsymbol{r},\boldsymbol{t})\propto \prod_{e=1}^E\Big(f_{\tau}(\tau_{e}; \mu_{s_e e}, V(\mu))\times \prod_{i\neq s_e}\big(1-F_{\tau}(\tau_{e}; \mu_{ie}, V(\mu)) \big)\Big),

@@

	where $f_{\tau}(\tau_{e}; \mu_{s_e e}, V(\mu))$ is the probability that the $e^{\textrm{th}}$ observed time increment comes from the specified distribution $f_\tau(\cdot)$ with the observed sender's mean $\mu_{s_e e}$, and $\prod_{i\neq s_e}\big(1-F_{\tau}(\tau_{e}; \mu_{ie},V(\mu)) \big)$ is the probability that the rest of (unobserved) senders for event $e$ all draw time increments greater than $\tau_e$. Moreover, under the existence of tied events, the conditional posterior of $\boldsymbol{\eta}$ is written as proportional to

@@

			\Pr(\boldsymbol{\eta}|\, \boldsymbol{u}, \boldsymbol{y},\boldsymbol{s}, \boldsymbol{r},\boldsymbol{t})&\propto \prod_{m=1}^M\Big(\prod_{e:t_e=t_m^*}f_{\tau}(t_m^*-t_{m-1}^*; \mu_{s_e e}, V(\mu)) \\&\times \prod_{i \notin \{s_e\}_{e:t_e=t_m^*}}\big(1-F_{\tau}(t_m^*-t_{m-1}^*; \mu_{ie}, V(\mu)) \big)\Big),

@@

	where $t_1^*,\ldots,t_M^*$ are the unique timepoints across $E$ events ($M \leq E$). If $M=E$ (i.e., no tied events), equation (\ref{eqn:latenttime2}) reduces to equation (\ref{eqn:latenttime}). Note that when we have the latent variable to quantify the variance in time increments $V(\mu)$ (based on the choice of timestamp distribution in Section \ref{subsec:Time}), we also use equation (\ref{eqn:latenttime}) (or equation (\ref{eqn:latenttime2}) in case there exist tied events) for the additional MH update---e.g., $\Pr(k|\, \boldsymbol{\eta},\boldsymbol{u}, \boldsymbol{y},\boldsymbol{s}, \boldsymbol{r},\boldsymbol{t})$ for Weibull, $\Pr(\theta|\, \boldsymbol{\eta},\boldsymbol{u}, \boldsymbol{y},\boldsymbol{s}, \boldsymbol{r},\boldsymbol{t})$ for gamma, and  $\Pr(\sigma^2_\tau| \,\boldsymbol{\eta},\boldsymbol{u}, \boldsymbol{y},\boldsymbol{s}, \boldsymbol{r},\boldsymbol{t})$ for log-normal.~\looseness=-1

@@

				\FOR{$i = 1$ to $A$}
				\FOR{$j = 1$ to $A$ ($j \neq i$)}

@@

				\IF {extra parameter for $V(\mu)$} 

@@

	we formulate our model specification through definitions of the receiver selection features $\boldsymbol{x}$ and event timing features $\boldsymbol{y}$. We then report a suite of experiments---out-of-sample prediction for model selection and posterior predictive checks---that illustrate how alternative formulations of the HEM can be compared, and evaluate how well our model recovers the distribution of the observed data. Finally, we demonstrate an exploratory analysis of Montgomery County email data using the model estimates to discover substantively meaningful patterns in organizational communication networks.

@@

		\item[7.] $\mbox{hyperedge}\_\,\mbox{size}(i)$: ${x}_{iej7}=\sum_{e^\prime: t_{e^\prime} \in l_e} \sum_{j=1}^A I(s_{e^\prime} = i)\,I(u_{e^\prime j} = 1)$;

@@

		\caption {Comparison of predictive performance between log-normal and exponential distributions: (a) correct sender posterior probability from sender predictions, (b) $F_1$ scores from receiver predictions, and (c) median absolute relative error from timestamp predictions. Blue line in (a) represents the correct sender probability expected by random guess---i.e., $1/A=1/18\approx0.056$.}

@@

			&\mu_{ie} = \eta_{1}+\eta_{2} y_{ie2}+\ldots+\eta_{7}y_{ie7}.

>>>>

	The hyperedge event model (HEM) specifies a generative process for $E$ unique hyperedge events that occur between $V$ number of nodes. A single hyperedge event is indexed by $e \in [E]$---where $[E]$ denotes a categorical set with $E$ levels $[E] = \{1,\ldots,E\}$---and consists of three components: the sender $s_e \in [V]$, an indicator vector of receivers $\boldsymbol{r}_e$---where $r_{ej}=1$ if $j \in [V]$ is a receiver of hyperedge event $e$ and 0 otherwise---and the timestamp $t_e \in (0, \infty)$. For simplicity, we assume that events are ordered by time such that $t_e \leq t_{e+1}$. While the model can be applied to two types of hyperedge events---events involving (1) one sender and one or more receivers, and (2) one or more senders and one receiver---here we only present the generative process for those involving one sender and one or more receivers (i.e., multicast). One notable feature of our generative process is that we draw auxiliary variables that serve as candidate data. Data is generated from the HEM through a sampling process applied to the auxiliary variables. The auxiliary variables drawn for event $e$ include, for each possible sender $i \in [V]$, a time increment from event $e-1$ at which sender $i$ would create event $e$, and an $V-1$ length vector indicating which nodes would be the receivers of event $e$ if it were directed by sender $i$.  The data generated for event $e$ under the HEM corresponds to the sender that would create event $e$ the soonest---at the smallest time increment from event $e-1$. The receivers of event $e$ generated under the HEM correspond to those receivers toward which the sender with the minimum time increment would have directed event $e$. We explain these steps in more detail below. Figure \ref{figure:diagram} presents an illustrative example on how the $e^{\textrm{th}}$ event is generated, assuming $t_{e-1} = 0$ and $V=5$. For hyperedge events that involve one receiver and one or more senders, we treat $s_e$ to be an indicator vector of senders $\boldsymbol{s}_e$ and $r_e$ to be the single receiver, and then follow the alternative generative process provided in Appendix A, which we derive as a simple reversal of the process used for multicasts (i.e., one sender and multiple receivers).~\looseness=-1

@@

	The HEM assumes that the sender of each hyperedge event is the sender that would initiate their respective event with the greatest urgency (i.e., the earliest timestamp). Our model thus assumes that for every event $e$, every possible sender $i$ generates a candidate receiver set that would be the receiver set of event $e$ if sender $i$ were the sender. For an event $e$, we first define an $V\times V$ matrix $\boldsymbol{u}_e$ where the $i^{\textrm{th}}$ row denotes sender $i$'s receiver vector of zeros and 1's---i.e., 1's indicate the nodes to which $i$ intends to direct event $e$. We then assume that each receiver vector $\boldsymbol{u}_{ie}$ comes from a modification of the multivariate Bernoulli (MB) distribution \citep{dai2013multivariate}---a model that has been used to model graphs in which the state of each edge indicator is drawn independently from an edge-specific Bernoulli distribution. Motivated by the Gibbs measure \citep{fellows2017removing}, we define a probability measure ``MB$_{G}$" in order to avoid drawing hyperedge events with no receiver. This probability measure amounts to a non-empty Gibbs measure, in which the all-zero vector is excluded from the support of the multivariate Bernoulli distribution. As a result, this measure helps us to (1) allow a sender to select multiple receivers for a single event, (2) force the sender to select at least one receiver, and (3) ensure a tractable normalizing constant for the receiver selection distribution. To be specific, we draw a binary vector $\boldsymbol{u}_{ie}= (u_{ie1},
	\ldots, u_{ieV})$ for $i$th row of matrix $\boldsymbol{u}$ using

@@

	\ldots, \lambda_{ieV})$ and $\lambda_{iej}$ is the linear combination of statistics relevant to the receiver selection process for every possible sender--receiver pair $(i,j)$ where $i \!\neq\! j$. In other words, we define

@@

	In particular, we define $\mbox{MB}_{G}(\boldsymbol{\lambda}_{ie})$ as
	\begin{equation}
		\begin{aligned}
			&\Pr(\boldsymbol{u}_{ie}|\,\boldsymbol{b}, \boldsymbol{x}_{ie}) = \frac{1}{C(\boldsymbol{\lambda}_{ie})}\exp\Big(\mbox{log}\big(\text{I}( \lVert \boldsymbol{u}_{ie}\rVert_1 > 0 )\big) + \sum_{j\neq i} \lambda_{iej}u_{iej}\Big) ,
		\end{aligned}
		\label{eqn:Gibbs}
	\end{equation}
	where $C(\boldsymbol{\lambda}_{ie})$ is the normalizing constant, $\lVert \cdot \rVert_1$ is the $\ell_1$-norm, and the log-indicator term $\mbox{log}\big(\text{I}( \lVert \boldsymbol{u}_{ie}\rVert_1 > 0 )\big)$ ensures that empty receiver sets are excluded from the distribution's support. These modeling assumptions facilitate efficient posterior inference since we can derive a closed form expression for the normalizing constant---i.e., $C(\boldsymbol{\lambda}_{ie})= \prod_{j\neq i} \big(\mbox{exp}(\lambda_{iej}) + 1\big)-1$---and thus do not need to perform brute-force summation over the support of $\boldsymbol{u}_{ie} \in \{0,1\}^V$. We provide detailed derivation steps for the normalizing constant $C(\boldsymbol{\lambda}_{ie})$ in Appendix B.

	% $C(\boldsymbol{\lambda}_{ie})= \prod_{j\neq i} \big(\mbox{exp}(\lambda_{iej}) + 1\big)-1$

	\subsection{Candidate timestamps}\label{subsec:Time}	
	 In modeling ``when," we do not directly model the timestamp $t_e$. Instead, we assume that each sender's ``time increment"---i.e., waiting time to next event since $t_{e-1}$---is drawn from a specific exponential family distribution. We assume that every sender $i \in [V]$ generates his/her own time increment from event $e-1$ to event $e$, which we define as $\tau_{ie}$. We further specify that the distribution of $\tau_{ie}$ has sender-specific mean $\mu_{ie}$ and common variance $\sigma_\tau^2$. Following the generalized linear model (GLM) framework \citep{nelder1972generalized}, we assume that sender $i$'s time increment $\tau_{ie} > 0$ satisfy~\looseness=-1
	\begin{equation}
		\begin{aligned}
\mbox{E}(\tau_{ie}) &= \mu_{ie},\\
\mbox{Var}(\tau_{ie}) &= \sigma_\tau^2,
		\end{aligned}
	\end{equation}
		% \footnote{The log-normal distribution is not exponential family but can be used via modeling of $\log(\tau_e)$.} 
	where the timing rate $\mu_{ie}$ for sender $i$ and event $e$ is determined by the linear combination of statistics relevant to timing selection process, similar to Section \ref{subsec: Tie}. Specifically, we define
	\begin{equation}
		\mu_{ie} = g^{-1}(\boldsymbol{\eta}^\top \boldsymbol{z}_{ie}),
	\end{equation}
	where $\boldsymbol{\eta}$ is a $Q$-dimensional vector of coefficients with a Normal prior $\boldsymbol{\eta} \sim N(\boldsymbol{\mu}_\eta,\Sigma_\eta)$, $\boldsymbol{z}_{ad}$ is a set of event timing features---covariates that could affect timestamps of events, and $g(\cdot)$ is the appropriate link function such as identity, log, or inverse. Possible choices of distribution include exponential, Weibull, gamma, and log-normal distributions, which are commonly used in time-to-event modeling \citep{rao2000applied,rizopoulos2012joint}. Based on the specific distribution, we may need another latent variable to account for the variance of time increments, such as the shape parameter for Weibull and gamma, or the variance parameter for log-normal, which is commonly used across $i\in [V]$ and $e \in [E]$. We use $f_\tau(\cdot; \mu, \sigma_\tau^2)$ and $F_\tau(\cdot; \mu, \sigma_\tau^2)$ to denote the probability density function (p.d.f) and cumulative density function (c.d.f), respectively, with mean $\mu$ and variance $\sigma_\tau^2$. 
	\subsection{Senders, receivers, and timestamps}\label{subsec:Observed}
		After the HEM draws the candidate timestamp vector $\boldsymbol{\tau}_e=\{\tau_{1e},\ldots,\tau_{Ve}\}$ at which the event would be created given the candidate sender and receiver combinations, finally our model assumes that the observed sender, receivers, and timestamp of hyperedge event $e$ are generated by selecting the sender--receiver-set pair with the smallest time increment:
			\begin{algorithm}[!t]
				\spacingset{1}
				\SetAlgoLined
				\caption{Generative process: one sender and one or more receivers}
				\begin{algorithmic}
					\STATE \textbf{Input}: number of events and nodes $(E, V)$, covariates $(\boldsymbol{x}, \boldsymbol{z})$, and coefficients $(\boldsymbol{b}, \boldsymbol{\eta})$
					\vskip 0.1in
					\FOR{$e=1$ to $E$}
					\FOR{$i=1$ to $V$}
					\FOR{$j=1$ to $V$ ($j \neq i$)}
					\STATE	set $\lambda_{iej} = {\boldsymbol{b}}^{\top}\boldsymbol{x}_{iej}$
					\ENDFOR
					\STATE	draw $\boldsymbol{u}_{ie}  \sim
					\mbox{MB}_G(\boldsymbol{\lambda}_{ie})$
					\STATE		set $\mu_{ie} = g^{-1}(\boldsymbol{\eta}^\top \boldsymbol{z}_{ie})$
					\STATE		draw $\tau_{ie} \sim f_\tau(\mu_{ie}, \sigma_\tau^2)$
					\ENDFOR

					\IF{$n \geq 2$ tied events} 
					\STATE	set $s_e,\ldots, s_{e+n-1}=\mbox{argmin}_{i}(\tau_{ie}),$
					\STATE	set $\boldsymbol{r}_e=\boldsymbol{u}_{s_e e},\ldots,\boldsymbol{r}_{e+n-1}=\boldsymbol{u}_{s_{e+n-1} e}$
					\STATE	set $t_e, \ldots, t_{e+n-1}=t_{e-1} + \min_i\tau_{ie}$
					\STATE		jump to $e = e+n$
					\ELSE
					\STATE	set $s_e= \mbox{argmin}_{i}(\tau_{ie})$
					\STATE		set $\boldsymbol{r}_e = \boldsymbol{u}_{s_e e}$
					\STATE	set $t_e =t_{e-1} + \min_i\tau_{ie}$
					\ENDIF
					\ENDFOR
				\end{algorithmic}
				\label{alg:generative}
			\end{algorithm}
	\begin{equation}
		\begin{aligned}
			s_e &= \mbox{argmin}_{i}(\tau_{ie}),\\
			\boldsymbol{r}_e &= \boldsymbol{u}_{s_e e},\\
			t_e &=t_{e-1} + \tau_{s_e e}.
		\end{aligned}
	\end{equation}
		Therefore, the HEM assumes a sender-driven process---i.e., the receivers and timestamp of an event are jointly determined by the sender's urgency to direct the event to those selected receivers---which is similar to \cite{snijders1996stochastic} in that each actor gets his/her own rate function to determine ``who will make the next change in network." Note that our generative process allows for tied events. In the case of tied events---i.e., multiple senders draw exactly the same candidate timestamps---we assume that all events are generated and occur simultaneously. Algorithm \ref{alg:generative} summarizes the entire generative process for hyperedge events with one sender and one or more receivers.~\looseness=-1

	\section{Posterior inference}\label{sec:inference}
	In this section we describe how we invert the generative process to obtain the posterior distribution over the latent variables---candidate receivers $\{\boldsymbol{u}_e\}_{e=1}^E$, coefficients for receiver selection features $\boldsymbol{b}$, and coefficients for event timing features $\boldsymbol{\eta}$---conditioned on the observed data $\{(s_e, \boldsymbol{r}_e, t_e)\}_{e=1}^E$, covariates $\{(\boldsymbol{x}_e, \boldsymbol{z}_e)\}_{e=1}^E$, and hyperparameters $(\boldsymbol{\mu}_b, \Sigma_b, \boldsymbol{\mu}_\eta, \Sigma_\eta)$. We draw the samples using Markov chain Monte Carlo (MCMC) methods, repeatedly resampling the value of each latent variable from its conditional posterior via a Metropolis-within-Gibbs sampling algorithm. In the next subsection, we provide each latent variable's conditional posterior along with pseudocode of MCMC in Algorithm \ref{alg:MCMC}. We also evaluate the correctness of both our mathematical derivations and software implemenation using the prior--posterior simulator test of \cite{geweke2004getting} in Appendix C.

@@

	In our model, direct computation of the posterior densities for the latent variables $\boldsymbol{b}$ and $\boldsymbol{\eta}$---i.e., $\Pr(\boldsymbol{b}|\,\boldsymbol{x},\boldsymbol{s}, \boldsymbol{r},\boldsymbol{t})$ and $\Pr(\boldsymbol{\eta}|\,\boldsymbol{z},\boldsymbol{s}, \boldsymbol{r},\boldsymbol{t})$---is not possible. However, it is possible to augment the data by candidate receivers $\boldsymbol{u}$ such that we can obtain their conditional posterior by conditioning on samples of $\boldsymbol{u}$. %from $P(\boldsymbol{b}, \boldsymbol{u}|\,\boldsymbol{x},\boldsymbol{a}, \boldsymbol{r},\boldsymbol{t})$ and $P(\boldsymbol{\eta}, \boldsymbol{u}|\, \boldsymbol{z},\boldsymbol{a}, \boldsymbol{r},\boldsymbol{t})$.

@@

		\prod_{i=1}^V \frac{1}{C(\boldsymbol{\lambda}_{ie})}\exp\Big(\mbox{log}\big(\text{I}( \lVert \boldsymbol{u}_{ie}\rVert_1 > 0)\big) + \sum\limits_{j \neq i} \lambda_{iej}u_{iej}\Big).

@@

		\Pr(\boldsymbol{\eta}|\, \boldsymbol{u}, \boldsymbol{z},\boldsymbol{s}, \boldsymbol{r},\boldsymbol{t})\propto \prod_{e=1}^E\Big(f_{\tau}(\tau_{e}; \mu_{s_e e}, \sigma_\tau^2)\times \prod_{i\neq s_e}\big(1-F_{\tau}(\tau_{e}; \mu_{ie}, \sigma_\tau^2) \big)\Big),

@@

	where $f_{\tau}(\tau_{e}; \mu_{s_e e}, \sigma_\tau^2)$ is the probability that the $e^{\textrm{th}}$ observed time increment comes from the specified distribution $f_\tau(\cdot)$ with the observed sender's mean $\mu_{s_e e}$, and $\prod_{i\neq s_e}\big(1-F_{\tau}(\tau_{e}; \mu_{ie},\sigma_\tau^2) \big)$ is the probability that the rest of (unobserved) senders for event $e$ all draw time increments greater than $\tau_e$. Moreover, under the existence of tied events, the conditional posterior of $\boldsymbol{\eta}$ is written as proportional to

@@

			\Pr(\boldsymbol{\eta}|\, \boldsymbol{u}, \boldsymbol{z},\boldsymbol{s}, \boldsymbol{r},\boldsymbol{t})&\propto \prod_{m=1}^M\Big(\prod_{e:t_e=t_m^*}f_{\tau}(t_m^*-t_{m-1}^*; \mu_{s_e e}, \sigma_\tau^2) \\&\times \prod_{i \notin \{s_e\}_{e:t_e=t_m^*}}\big(1-F_{\tau}(t_m^*-t_{m-1}^*; \mu_{ie}, \sigma_\tau^2) \big)\Big),

@@

	where $t_1^*,\ldots,t_M^*$ are the unique timepoints across $E$ events ($M \leq E$). If $M=E$ (i.e., no tied events), equation (\ref{eqn:latenttime2}) reduces to equation (\ref{eqn:latenttime}). Note that when we have the latent variable to quantify the variance in time increments $\sigma_\tau^2$ (based on the choice of timestamp distribution in Section \ref{subsec:Time}), we also use equation (\ref{eqn:latenttime}) (or equation (\ref{eqn:latenttime2}) in case there exist tied events) for the additional MH update---e.g., $\Pr(k|\, \boldsymbol{\eta},\boldsymbol{u}, \boldsymbol{z},\boldsymbol{s}, \boldsymbol{r},\boldsymbol{t})$ for Weibull, $\Pr(\theta|\, \boldsymbol{\eta},\boldsymbol{u}, \boldsymbol{z},\boldsymbol{s}, \boldsymbol{r},\boldsymbol{t})$ for gamma, and  $\Pr(\sigma^2_\tau| \,\boldsymbol{\eta},\boldsymbol{u}, \boldsymbol{z},\boldsymbol{s}, \boldsymbol{r},\boldsymbol{t})$ for log-normal.~\looseness=-1

@@

				\FOR{$i = 1$ to $V$}
				\FOR{$j = 1$ to $V$ ($j \neq i$)}

@@

				\IF {extra parameter for $\sigma_\tau^2$}

@@

	we formulate our model specification through definitions of the receiver selection features $\boldsymbol{x}$ and event timing features $\boldsymbol{z}$. We then report a suite of experiments---out-of-sample prediction for model selection and posterior predictive checks---that illustrate how alternative formulations of the HEM can be compared, and evaluate how well our model recovers the distribution of the observed data. Finally, we demonstrate an exploratory analysis of Montgomery County email data using the model estimates to discover substantively meaningful patterns in organizational communication networks.

@@

		\item[7.] $\mbox{hyperedge}\_\,\mbox{size}(i)$: ${x}_{iej7}=\sum_{e^\prime: t_{e^\prime} \in l_e} \sum_{j=1}^V I(s_{e^\prime} = i)\,I(u_{e^\prime j} = 1)$;

@@

		\caption {Comparison of predictive performance between log-normal and exponential distributions: (a) correct sender posterior probability from sender predictions, (b) $F_1$ scores from receiver predictions, and (c) median absolute relative error from timestamp predictions. Blue line in (a) represents the correct sender probability expected by random guess---i.e., $1/V=1/18\approx0.056$.}

@@

			&\mu_{ie} = \eta_{1}+\eta_{2} z_{ie2}+\ldots+\eta_{7}z_{ie7}.
